---
title: "K-Means Demo"
author: "Yadder Aceituno"
date: "December 25, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6, fig.height = 3)

library(ggplot2);
library(plyr);
```

## Summary

Clustering is one of the most data exploratory analysis used to get a insights from data. It helps to identify subgroups that can help us to identify how the data is structured.

Kmeans is an algorithm that tries to split the dataset into K pre-defined distinct groups (non-overlaping), which are called clusters.

Using the application we can create a random dataset and then running the k-means algorithm to see how the data is clusterized. We can visualize the plot to learn how the different clusters, its centroids and its points are created.

## Configuration Panel

To build the plot, we have the folowing options:

- Slider for Number of Points: to define how many points will be generated randomly.
- Button to generate random data: to generate 20 random points.
- Slider for Number of Clusters: to define how many clusters to show.
- Checkbox for configuration: to enable/disable centroids and areas.
- Button to build clusters: to build the clusters with k-mean algorithm.

## Random DataSet

We can create a random dataset with 20 (default) or more points by pressing the **Generate Random Points** button. It will execute the next code:

```{r random_code, echo = TRUE, eval=FALSE}
qtyPoints <- input$slidePoints;
dfData <- data.frame( x = runif(qtyPoints, min = 0, max = 100), 
                    y = runif(qtyPoints, min = 0, max = 100));
```

```{r random_plot, echo = FALSE}
qtyPoints <- 20;
dfData <- data.frame( x = runif(qtyPoints, min = 0, max = 100), 
                    y = runif(qtyPoints, min = 0, max = 100));

myPlot <- ggplot(mapping = aes(x = x, y = y));
myPlot <- myPlot + geom_point(dfData, size = 3.5, mapping = aes());

# Setting axis limits
myPlot <-  myPlot +
  theme(legend.position = "none") +
  xlim(0, 105) +
  ylim(0, 105);

myPlot;
```

## Making Clusters

By pressing **Make Clusters** button, it will execute the kmeans algorithm on the data and then will plot the clusters (according configuration).

```{r kmeans_code, echo=TRUE, eval=FALSE}
objCluster <- kmeans(dfData, centers = input$slideClusters);
```

```{r kmeans_plot, echo=FALSE}
qtyClusters <- 3

objCluster <- kmeans(dfData, centers = qtyClusters)

colors <- c("#550000", "#D46A6A",
              "#FFD1AA", "#804515",
              "#116611", "#88CC88",
              "#003333", "#407F7F",
              "#D4D369", "#555500");

myPlot <- ggplot(mapping = aes(x = x, y = y));

# Setting points
myPlot <- myPlot + geom_point(dfData, size = 3, mapping = aes(color = as.factor(objCluster$cluster))) +
        scale_colour_manual(values = colors[1:qtyClusters]);

dfData$cluster <- objCluster$cluster;

# Calculating areas
fncFindHull <- function(dfTemp) dfTemp[chull(dfTemp$x, dfTemp$y), ];
hulls <- ddply(dfData, "cluster", fncFindHull);

# Plotting areas.
myPlot <- myPlot + geom_polygon(data = hulls, alpha = 0.5, aes(colour = as.factor(cluster), fill = as.factor(cluster))) + 
  scale_fill_manual(values = colors[1:qtyClusters]);

# Plotting centers
myPlot <- myPlot + geom_point(data.frame(objCluster$centers), shape = 4, size = 3.5, mapping = aes());

# Setting axis limits
myPlot <-  myPlot +
      theme(legend.position = "none") +
      xlim(0, 105) +
      ylim(0, 105);

myPlot;

```

